---
title: "R Notebook"
output:
  html_document: default
  html_notebook: default
---


This is a notebook for the examination of CDR3s in the repertoire of peanut allergic individuals

Load in packages and the raw data
```{r}
library(data.table)
library(ggplot2)
library(gridExtra)
library(magrittr)
library(dplyr)

switch(Sys.info()[['user']],
       nealp = {file.path <- "C:/Users/nealp/Dropbox (Personal)/Extension School/Thesis/figures"
       raw.file.path <- "C:/Users/nealp/Dropbox (Partners HealthCare)/Projects/PNOIT2-1037/TCRB sequencing and HLA typing data"},
       wayne1 = {file.path <- "~/Dropbox (Personal)/Extension School/Thesis/figures"
       raw.file.path <- "~/Dropbox (Partners HealthCare)/Projects/PNOIT2-1037/TCRB sequencing and HLA typing data"},
        stop("I don't recognize your username, type Sys.info() to find out what it is.")
)
source("neals_tcr_functions.R")

# Load in the raw data
load(paste(raw.file.path, "neals.thesis.data/parsed.data.baseline.rda", sep = "/"))
load(paste(raw.file.path, "neals.thesis.data/enriched.CDR3s.rda", sep = "/"))

# Make a dataframe for the enriched CDR3s
enriched.CDR3.df <- do.call(rbind, enriched.CDR3s)

```

First we wanted to look at overall sequences in our dataset.  
```{r}
# Create a dataframe with general information
info.df <- data.frame(id = names(enriched.CDR3s),
                resting.CDR3nuc = unlist(lapply(data.parse[grep("neg", names(data.parse))], nrow)),
                resting.CDR3AA = unlist(lapply(data.parse[grep("neg", names(data.parse))], function(x){
                  length(unique(x$CDR3.amino.acid.sequence))})),
                activated.CDR3nuc = unlist(lapply(data.parse[grep("pos", names(data.parse))], nrow)),
                activated.CDR3AA = unlist(lapply(data.parse[grep("pos", names(data.parse))], function(x){
                  length(unique(x$CDR3.amino.acid.sequence))
                })),
                psCDR3nuc = unlist(lapply(grep("pos", names(data.parse), value = TRUE), function(x){
                  nrow(data.parse[[x]][data.parse[[x]]$CDR3.amino.acid.sequence %in% 
                                         enriched.CDR3s[[gsub("([0-9]+).*$", "\\1", x)]]$CDR3.amino.acid.sequence,])
                })),
                psCDR3AA = unlist(lapply(enriched.CDR3s, nrow))
                )
info.df$psCDR3nuc.perc.act <- (info.df$psCDR3nuc / info.df$activated.CDR3nuc) * 100
info.df$psCDR3AA.perc.act <- (info.df$psCDR3AA / info.df$activated.CDR3AA) * 100

plot.df <- melt(info.df, id.vars = "id")

plt.1 <- ggplot(plot.df[grep("resting", plot.df$variable),], aes(x = variable, y = value)) +
  geom_boxplot() +
  ylab("sequences") + xlab("") +
  scale_x_discrete(labels = c("CDR3 nucleotide sequences", "CDR3 AA sequences")) +
  ggtitle("Resting CDR3s") +
  theme_bw()
plt.2 <- ggplot(plot.df[grep("activated", plot.df$variable),], aes(x = variable, y = value)) +
  geom_boxplot() +
  ylab("sequences") + xlab("") +
  scale_x_discrete(labels = c("CDR3 nucleotide sequences", "CDR3 AA sequences")) +
  ggtitle("activated CDR3s") +
  theme_bw()
grid.arrange(plt.1, plt.2, ncol = 2)

ggplot(plot.df[plot.df$variable == "psCDR3nuc" | plot.df$variable == "psCDR3AA",], aes(x = variable, y = value)) + geom_boxplot() +
  ylab("sequences") + xlab("") +
  scale_x_discrete(labels = c("CDR3 nucleotide sequences", "CDR3 AA sequences")) +
  ggtitle("psCDR3s") +
  theme_bw(base_size = 15)
ggplot(plot.df[grep("perc.act", plot.df$variable),], aes(x = variable, y = value)) + geom_boxplot() +
  ylab("percent of activated CDR3s") + xlab("") +
  scale_x_discrete(labels = c("CDR3 nucleotide sequences", "CDR3 AA sequences")) +
  ggtitle("percent of activated CDR3s that are psCDR3s") +
  theme_bw(base_size = 20)

```

Here we can see that our statistical enrichment selects only ~15% of the total activated compartment, indicating the stringency with which it selects sequences


```{r hamming}
# Perform hamming on entire list of peanut CDR3s
min.ham <- CDR3_dist(enriched.CDR3.df$CDR3.amino.acid.sequence, method = "hamming")
min.lev <- CDR3_dist(enriched.CDR3.df$CDR3.amino.acid.sequence, method = "lv")
# Create a list of all of the CDR3s from CD154- to be used as a control
rest.CDR3s <- lapply(data.parse[grep("neg",names(data.parse))], function(x){
  x <- unique(x$CDR3.amino.acid.sequence)
}) %>% unlist(., use.names = FALSE)

# Get hamming distance on all CD154+ and CD154-
neg.ham.df <- data.frame(distance = 0:20)
neg.lev.df <- data.frame(distance = 0:20)

for(i in 1:100){
  sample <- sample(rest.CDR3s, nrow(enriched.CDR3.df))
  neg.ham <- CDR3_dist(sample, "hamming")
  neg.lev <- CDR3_dist(sample, "lv")
  
  neg.ham.df <- left_join(neg.ham.df, neg.ham, by = "distance", all = TRUE)
  neg.lev.df <- left_join(neg.lev.df, neg.lev, by = "distance", all = TRUE)
}

# Clean up negative dataframes, add median percent and standard deviation
neg.ham.df[is.na(neg.ham.df)] <- 0
neg.ham.df$median <- apply(neg.ham.df[2:ncol(neg.ham.df)], 1, median)
neg.ham.df$sd <- apply(neg.ham.df[2:ncol(neg.ham.df)], 1, sd)

# Clean up negative dataframes, add median percent and standard deviation
neg.lev.df[is.na(neg.lev.df)] <- 0
neg.lev.df$median <- apply(neg.lev.df[2:ncol(neg.lev.df)], 1, median)
neg.lev.df$sd <- apply(neg.lev.df[2:ncol(neg.lev.df)], 1, sd)

# Create a list of all of the CDR3s from CD154- to be used as a control
act.CDR3s <- lapply(data.parse[grep("pos",names(data.parse))], function(x){
  x <- unique(x$CDR3.amino.acid.sequence)
}) %>% unlist(., use.names = FALSE)


# Do the same for all CD154+
pos.all.ham.df <- data.frame(distance = 0:20)
pos.all.lev.df <- data.frame(distance = 0:20)
for(i in 1:100){
  sample = sample(act.CDR3s, nrow(enriched.CDR3.df))
  pos.all.ham <- CDR3_dist(sample, method = "hamming")
  pos.all.lev <- CDR3_dist(sample, method = "lv")
  
  pos.all.ham.df <- left_join(pos.all.ham.df, pos.all.ham, by = "distance", all = TRUE) 
  pos.all.lev.df <- left_join(pos.all.lev.df, pos.all.lev, by = "distance", all = TRUE)
}

# Clean up negative dataframe, add median percent and standard deviation
pos.all.ham.df[is.na(pos.all.ham.df)] <- 0
pos.all.ham.df$median <- apply(pos.all.ham.df[2:ncol(pos.all.ham.df)], 1, median)
pos.all.ham.df$sd <- apply(pos.all.ham.df[2:ncol(pos.all.ham.df)], 1, sd)

# Clean up negative dataframe, add median percent and standard deviation
pos.all.lev.df[is.na(pos.all.lev.df)] <- 0
pos.all.lev.df$median <- apply(pos.all.lev.df[2:ncol(pos.all.lev.df)], 1, median)
pos.all.lev.df$sd <- apply(pos.all.lev.df[2:ncol(pos.all.lev.df)], 1, sd)


sd.ham <- neg.ham.df$sd[as.numeric(as.character(neg.ham.df$distance)) <=15] %>%
  append(pos.all.ham.df$sd[as.numeric(as.character(pos.all.ham.df$distance)) <=15]) %>%
  append(rep(NA, 16))

sd.lev <- neg.lev.df$sd[as.numeric(as.character(neg.lev.df$distance)) <=15] %>%
  append(pos.all.lev.df$sd[as.numeric(as.character(pos.all.lev.df$distance)) <=15]) %>%
  append(rep(NA, 16))

min.ham <- full_join(min.ham, neg.ham.df, by = "distance") %>%
  select(c(1, 2, ncol(.) - 1)) %>%
  full_join(., pos.all.ham.df, by = "distance") %>%
  select(c(1,3, ncol(.) - 1, 2))
colnames(min.ham) <- c("distance", "CD154-", "CD154+", "enriched.CD154+")

min.lev <- full_join(min.lev, neg.lev.df, by = "distance") %>%
  select(c(1, 2, ncol(.) - 1)) %>%
  full_join(., pos.all.lev.df, by = "distance") %>%
  select(c(1,3, ncol(.) - 1, 2))
colnames(min.lev) <- c("distance", "CD154-", "CD154+", "enriched.CD154+")


min.ham[is.na(min.ham)] <- 0
min.ham <- min.ham[!min.ham$distance == "Inf",]
min.ham <- min.ham[order(min.ham$distance),]

min.ham.melt <- melt(min.ham, id.vars = "distance")
min.ham.melt$distance <- as.integer(min.ham.melt$distance)
min.ham.melt <- min.ham.melt[min.ham.melt$distance <=15,] %>%
  cbind(., sd)

min.lev[is.na(min.lev)] <- 0
min.lev <- min.lev[!min.lev$distance == "Inf",]
min.lev <- min.lev[order(min.lev$distance),]

min.lev.melt <- melt(min.lev, id.vars = "distance")
min.lev.melt$distance <- as.integer(min.lev.melt$distance)
min.lev.melt <- min.lev.melt[min.lev.melt$distance <=15,] %>%
  cbind(., sd)

ggplot(min.ham.melt, aes(x = distance, y = value, fill = variable, width = .80)) +
  geom_bar(stat = "identity", position = position_dodge()) + scale_x_continuous(breaks = c(0:15)) +
  scale_fill_manual(values = c('#E69F00', "#999999", "#b30000"), labels = c("CD154- T cells", "All CD154+ T cells",
                                                                            "Peanut Enriched CD154+ cells")) +
  geom_errorbar(aes(ymin = value, ymax = sd + value), width = .3, position = position_dodge(0.8)) +
  ylab("percent of total clones") + xlab("AA differences to next closest CDR3") +
  ggtitle("Hamming distance for peanut CDR3s")

```

Hamming 

Add a new chunk by clicking the *Insert Chunk* button on the toolbar or by pressing *Ctrl+Alt+I*.

When you save the notebook, an HTML file containing the code and output will be saved alongside it (click the *Preview* button or press *Ctrl+Shift+K* to preview the HTML file).
